---
title: "bios611_shiny"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages 

```{r}
library(tidyverse)
library(dplyr)
library(shiny)
library(ggplot2)
library(plotly)
```

## Load data 

COVID data 
Country demographic data - from 2017 

```{r}
covid <- read_csv("worldometer_coronavirus_summary_data.csv")
#country <- read_csv("UNdata_Export_2020.csv")
demographics <- read_csv("UNdata_Export_2017.csv")

colnames(demographics) <- c("country", "year", "area", "sex", "age", "record_type", "reliability", "source_year", "value", "value_footnotes")
```

## Tidy data 

Sorting data by number of data points associated with a given country shows us that there are a few singletons that don't appear to represent countries at all, so remove those:

```{r}
country_counts <- demographics %>% count(country) %>% filter(n > 1)
demographics <- demographics %>% filter(country %in% country_counts$country)
```

Remove non-ASCII characters and translate to camel case (to match up with covid data): 

```{r}
simplify_strings <- function(s){
    s %>% 
        str_to_title() %>%
        str_trim() %>%
        str_replace_all("-"," ") %>%
        str_replace_all(",","") %>% 
        str_replace("\\s*\\([^\\)]+\\)", "") %>% # remove anything in parentheses 
        iconv(to='ASCII//TRANSLIT') # remove accented characters 
}

demographics$country <- simplify_strings(demographics$country)
```

Perform the following filtering:
1. Filter out area-specific data (rural vs. urban)
2. Filter out sex-specific data
3. Final figures only
4. Age ranges only (filter out specific age data)

Select first row in each set of distinct combinations of country and age. 

```{r}
demographics <- demographics %>% filter(area == "Total", sex == "Both Sexes", reliability == "Final figure, complete", nchar(age) > 3) %>%
  distinct(country, age, .keep_all = TRUE) %>%
  select(country, age, value)
```

Pivot wide so that each column is the population for a specific age group: 

```{r}
demo_wide <- demographics %>% pivot_wider(id_cols = country, names_from = age, values_from = value)
```

Remove "unknown" column and redundant columns:

```{r}
demo_wide <- demo_wide %>% select(-Unknown, -'1 - 4', -'0 - 14', -'15 - 64', -'95 - 98', -'65 +', -'75 +', -'80 +', -'85 +', -'90 +',-'95 +', -'99 +', -'100 +')
```

## Calculate metrics 

Proportion of total population for each country that is over 40, 50, 60, 70, 80 

```{r}
demo_wide <- demo_wide %>% rowwise() %>% 
  mutate(over_40 = sum(c_across('40 - 44':'110 +'), na.rm = TRUE)/Total) %>%
  mutate(over_50 = sum(c_across('50 - 54':'110 +'), na.rm = TRUE)/Total) %>%
  mutate(over_60 = sum(c_across('60 - 64':'110 +'), na.rm = TRUE)/Total) %>%
  mutate(over_70 = sum(c_across('70 - 74':'110 +'), na.rm = TRUE)/Total) %>%
  mutate(over_80 = sum(c_across('80 - 84':'110 +'), na.rm = TRUE)/Total) %>%
  ungroup()
```

## Join country data with covid data 

Do an inner join, so the resulting table has countries which we have both demographic and covid data for. 

```{r}
combined <- inner_join(demo_wide, covid, by = "country")
```

Which countries from the demographics data set weren't included in the join? 
```{r}
demo_wide$country[which(!(demo_wide$country %in% combined$country))]
```

Rename countries in demographics data set that don't match up with country names in COVID-19 data set (didn't bother looking the other way, because there are several countries in the COVID-19 dataset for which we wont' have demographic data): 

```{r}
demo_wide$country[which(demo_wide$country == "Czechia")] <- "Czech Republic"
demo_wide$country[which(demo_wide$country == "Faroe Islands")] <- "Faeroe Islands"
demo_wide$country[which(demo_wide$country == "North Macedonia")] <- "Macedonia"
demo_wide$country[which(demo_wide$country == "Republic Of Moldova")] <- "Moldova"
demo_wide$country[which(demo_wide$country == "Republic Of South Sudan")] <- "Sudan"
demo_wide$country[which(demo_wide$country == "United Kingdom Of Great Britain And Northern Ireland")] <- "UK"
demo_wide$country[which(demo_wide$country == "United States Of America")] <- "USA"
```

Re-do the join: 

```{r}
rm(combined)
combined <- inner_join(demo_wide, covid, by = "country")
```

## Create figures 

```{r}
ggplot(combined, aes(x = over_40, y = total_deaths_per_1m_population)) + geom_point()
```

```{r}
ggplot(combined, aes(x = over_70, y = total_deaths_per_1m_population)) + geom_point()
```

## Launch shiny app 

```{r}
shinyApp(
  ui = fluidPage(
    titlePanel("Country demographics vs. COVID-19 deaths by age"),
    
    sidebarLayout(
      sidebarPanel(
        varSelectInput(inputId="x_variable",
                       label="Variable (x):",
                       combined,
                       multiple=F),
        varSelectInput(inputId="y_variable",
                       label="Variable (y):",
                       combined,
                       multiple=F)
      ),
      mainPanel(
        plotlyOutput(outputId = "plot1")
      )
    )
  ),
  server = function(input, output) {
    output$plot1 <- renderPlotly({
      
      ggplot(combined, aes(!!input$x_variable, !!input$y_variable)) + geom_point()

    })
  }
)

shinyApp(ui=ui,server=server, options=list(port=8080, host="0.0.0.0"))
```




